---
title: "County Comparison plots"
author: "Jean Czerlinski Whitmore (jeanimal)"
date: "5/3/2020"
output:
  html_document:
    code_folding: hide
    keep_md: true
    number_sections: yes
    toc: yes
    toc_depth: 3
---

I noticed counties with protests had a double-hump shape.  I wanted to compare this shape with other counties in the same state.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
  fig.path = "county_compare_figs/county-"
)
```

```{r load-libraries, echo=FALSE}
library(tidyverse)
library(scales)
library(ggplot2)

source('covid_log_log_diff/functions.R')
```

## Loading data

```{r covid-cases-by-us-county}
covidByCounty <- loadCovidDataByGeo("US_COUNTY")
```

## Steep increase counties

Counties that ever hit 1,000 new cases per day.
```{r}
steepIncrease <- covidByCounty %>% dplyr::filter(newCasesPerDay > 1000) %>% dplyr::filter(state != "_ALL_")
steepIncreaseNames <- unique(steepIncrease$state)
steepIncreaseNames
```

```{r}

plotData <- covidByCounty %>%
  dplyr::filter(state %in% steepIncreaseNames)
```

```{r plot-steep-counties, fig.width=12, fig.height = 14}
ggplot(plotData, aes(x=cases, y=smoothed, group = state)) +
    geom_line(data = plotData %>% rename(group = state),
              aes(x = cases, y = smoothed, group = group), color = "grey") +
    geom_line(aes(y = smoothed), color = "black") +
    scale_x_log10(label = comma, breaks = c(100, 1000, 100000)) + 
    scale_y_log10(label = comma) +
    coord_equal() +
    labs(x = 'Total confirmed cases',
         y = 'New confirmed cases per day',
         title = 'Trajectory of COVID-19 cases in steep increase counties') +
    facet_wrap(~ state) +
    theme_minimal()
```
## New York

```{r}
plotData <- covidByCounty %>%
  dplyr::filter(us_state=="New York")
```

```{r plot-ny-counties, fig.width=9, fig.height = 8}
ggplot(plotData, aes(x=cases, y=smoothed, group = state)) +
    geom_line(data = plotData %>% rename(group = county),
              aes(x = cases, y = smoothed, group = group), color = "grey") +
    geom_line(aes(y = smoothed), color = "black") +
    scale_x_log10(label = comma, breaks = c(100, 1000, 100000)) + 
    scale_y_log10(label = comma) +
    coord_equal() +
    labs(x = 'Total confirmed cases',
         y = 'New confirmed cases per day',
         title = 'Trajectory of COVID-19 cases in New York counties') +
    facet_wrap(~ county) +
    theme_minimal()
```

## New Jersey

```{r}
plotData <- covidByCounty %>%
  dplyr::filter(us_state=="New Jersey")
```

```{r plot-nj-counties, fig.width=9, fig.height = 8}
ggplot(plotData, aes(x=cases, y=smoothed, group = state)) +
    geom_line(data = plotData %>% rename(group = county),
              aes(x = cases, y = smoothed, group = group), color = "grey") +
    geom_line(aes(y = smoothed), color = "black") +
    scale_x_log10(label = comma, breaks = c(100, 1000, 100000)) + 
    scale_y_log10(label = comma) +
    coord_equal() +
    labs(x = 'Total confirmed cases',
         y = 'New confirmed cases per day',
         title = paste('Trajectory of COVID-19 cases in', 'New Jersey', 'counties')) +
    facet_wrap(~ county) +
    theme_minimal()
```

## Michigan

```{r}
plotData <- covidByCounty %>%
  dplyr::filter(us_state=="Michigan")
```

```{r plot-mi-counties, fig.width=9, fig.height = 8}
ggplot(plotData, aes(x=cases, y=smoothed, group = state)) +
    geom_line(data = plotData %>% rename(group = county),
              aes(x = cases, y = smoothed, group = group), color = "grey") +
    geom_line(aes(y = smoothed), color = "black") +
    scale_x_log10(label = comma, breaks = c(100, 1000, 100000)) + 
    scale_y_log10(label = comma) +
    coord_equal() +
    labs(x = 'Total confirmed cases',
         y = 'New confirmed cases per day',
         title = 'Trajectory of COVID-19 cases in Michigan counties') +
    facet_wrap(~ county) +
    theme_minimal()
```


## Ohio

```{r}
plotData <- covidByCounty %>%
  dplyr::filter(us_state=="Ohio")
```

```{r plot-oh-counties, fig.width=9, fig.height = 8}
ggplot(plotData, aes(x=cases, y=smoothed, group = state)) +
    geom_line(data = plotData %>% rename(group = county),
              aes(x = cases, y = smoothed, group = group), color = "grey") +
    geom_line(aes(y = smoothed), color = "black") +
    scale_x_log10(label = comma, breaks = c(100, 1000, 100000)) + 
    scale_y_log10(label = comma) +
    coord_equal() +
    labs(x = 'Total confirmed cases',
         y = 'New confirmed cases per day',
         title = paste('Trajectory of COVID-19 cases in', 'Ohio', 'counties')) +
    facet_wrap(~ county) +
    theme_minimal()
```

## California

```{r}
plotData <- covidByCounty %>%
  dplyr::filter(us_state=="California")
```

```{r plot-ca-counties, fig.width=9, fig.height = 8}
ggplot(plotData, aes(x=cases, y=smoothed, group = state)) +
    geom_line(data = plotData %>% rename(group = county),
              aes(x = cases, y = smoothed, group = group), color = "grey") +
    geom_line(aes(y = smoothed), color = "black") +
    scale_x_log10(label = comma, breaks = c(100, 1000, 100000)) + 
    scale_y_log10(label = comma) +
    coord_equal() +
    labs(x = 'Total confirmed cases',
         y = 'New confirmed cases per day',
         title = 'Trajectory of COVID-19 cases in California counties') +
    facet_wrap(~ county) +
    theme_minimal()
```

## Louisiana

No protests against lockdown in Tennessee but I am curious how it looks.  I removed Bledsoe county because it seems to have a data error (100 million new cases on one day).

```{r}
plotData <- covidByCounty %>%
  dplyr::filter(us_state=="Louisiana")
```

```{r plot-louisiana-counties, fig.width=10, fig.height = 12}
ggplot(plotData, aes(x=cases, y=smoothed, group = state)) +
    geom_line(data = plotData %>% rename(group = county),
              aes(x = cases, y = smoothed, group = group), color = "grey") +
    geom_line(aes(y = smoothed), color = "black") +
    scale_x_log10(label = comma, breaks = c(100, 1000, 10000)) + 
    scale_y_log10(label = comma) +
    coord_equal() +
    labs(x = 'Total confirmed cases',
         y = 'New confirmed cases per day',
         title = 'Trajectory of COVID-19 cases in Louisiana counties') +
    facet_wrap(~ county) +
    theme_minimal()
```

## Tennessee

No protests against lockdown in Tennessee but I am curious how it looks.  I removed Bledsoe county because it seems to have a data error (100 million new cases on one day).

```{r}
plotData <- covidByCounty %>%
  dplyr::filter(us_state=="Tennessee") %>% dplyr::filter(county!="Bledsoe")
```

```{r plot-tn-counties, fig.width=6, fig.height = 8}
ggplot(plotData, aes(x=cases, y=smoothed, group = state)) +
    geom_line(data = plotData %>% rename(group = county),
              aes(x = cases, y = smoothed, group = group), color = "grey") +
    geom_line(aes(y = smoothed), color = "black") +
    scale_x_log10(label = comma, breaks = c(100, 1000, 10000)) + 
    scale_y_log10(label = comma) +
    coord_equal() +
    labs(x = 'Total confirmed cases',
         y = 'New confirmed cases per day',
         title = 'Trajectory of COVID-19 cases in Tennessee counties') +
    facet_wrap(~ county) +
    theme_minimal()
```