---
title: "State mandate plots"
author: "Jean Czerlinski Whitmore (jeanimal)"
date: "4/4/2020"
output:
  html_document:
    code_folding: hide
    keep_md: true
    number_sections: yes
    toc: yes
    toc_depth: 3
---

This combines data on state mandates with cases to help visualize the affect, if any, of these mandates on the growth of cases.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
  fig.path = "README_figs/README-"
)
```

```{r load-libraries, echo=FALSE}
library(tidyverse)
library(scales)
library(ggplot2)

source('covid_log_log_diff/functions.R')
```

## Loading data

```{r covid-cases-by-state}
covidByState <- loadAndFormatNytimesCovidPerState()
```

```{r mandates-by-state}
mandatesByState <- read.csv2('https://raw.githubusercontent.com/jeanimal/covid_log_log_diff/master/covid_log_log_diff/data/covid_state_mandates.csv', sep=",", stringsAsFactors=FALSE, na.strings=c(""))
mandatesByState$stayhome <- as.Date(mandatesByState$stayhome)
mandatesByState$schools <- as.Date(mandatesByState$schools)
mandatesByState$nonessential <- as.Date(mandatesByState$nonessential)
mandatesByState$travel <- as.Date(mandatesByState$travel)
```

## Munging data

```{r smooth-covid-by-state}
covidByState<-covidByState %>% 
  dplyr::filter(!is.na(newCasesPerDay), 
                !is.na(cases), 
                newCasesPerDay > 0, 
                cases > 0)  %>%
  dplyr::select(-prevDate,-prevCases)

# create loess-smoothed versions of time series for each state
covidByStateSmoothed <- covidByState %>%
  filter(!(state %in% c("USA","Northern Mariana Islands","Virgin Islands","Guam"))) %>%
  group_by(state) %>%
  do(data.frame(.,
                smoothed = 10^predict(loess(log10(newCasesPerDay) ~ log10(cases), data = .), .))) %>%
  ungroup()
```

## Plot helpers

```{r function-join-mandates-with-cases}
# joinMandatesWithCases joins mandateByState with the case data in covidByState
# It uses the state and date in covidByState for the state at mandateColName date.
# mandatesByState should have columns:
# - state
# - [mandateColName]
# covidByState should have columns:
# - state
# - date
# - whatever case info you want
joinMandatesWithCases <- function(mandatesByState, mandateColName, covidByState) {
  mandatesByStateWithCases <- mandatesByState[, c("state", mandateColName)]
  # The c() function messes up evaluation so I use a hacky renaming.
  names(mandatesByStateWithCases)[[2]] <- "date"
  mandatesByStateWithCases <- mandatesByStateWithCases %>% left_join(covidByState, by=c("state", "date"))
  names(mandatesByStateWithCases)[[2]] <- mandateColName
  mandatesByStateWithCases
}
```


```{r function-faceted-plot-with-mandates}
# Makes a faceted plot of case growth with a dot representing the start of a mandate.
# Both covidByStateSmoothed and mandatesByStateWithCases are data frames with columns
# - cases
# - smoothed
# - state
# However, while covidByStateSmoothed should have data for many dates,
# mandatesByStateWithCases should have data for only one date, the mandate start date.
# The final plot does *not* have dates on an axis, which is why the association between
# these data frames is indirect.
facetedPlotWithMandates <- function(covidByStateSmoothed, mandatesByStateWithCases, mandateName) {
  ggplot(covidByStateSmoothed, aes(x=cases, y=smoothed, group = state)) +
    geom_line(data = covidByStateSmoothed %>% rename(group = state),
              aes(x = cases, y = smoothed, group = group), color = "grey") +
    geom_line(aes(y = smoothed), color = "red") +
    geom_point(data=mandatesByStateWithCases, aes(x=cases, y=smoothed)) +
    scale_x_log10(label = comma, breaks = c(100, 1000, 100000)) + 
    scale_y_log10(label = comma) +
    coord_equal() +
    labs(x = 'Total confirmed cases',
         y = 'New confirmed cases per day',
         title = 'Trajectory of COVID-19 cases in the U.S.',
         subtitle = paste0('Dot is start of ', mandateName)) +
    facet_wrap(~ state) +
    theme_minimal()
}

```


## Plots

The dot for the start of a mandate is plotted by looking up the number of cases and new
cases on that day.

In general, we would expect a delay of 1-2 weeks between the date an order starts and a change in cases because the virus incubation period is at least that long.

## Temporary

Temporary: Remove states I do not have any mandate data for.
```{r}
background_states <- c("New York", "New Jersey", "California", "Michigan", "Louisiana", "Florida", "Massachusetts", "Illinois", "Pennsylvania", "Washington")
covidByStateSmoothed <- covidByStateSmoothed %>%
filter(state %in% c("Alabama", "Alaska", "Arizona", "Tennessee", "New York", background_states))
```


## The effect of Stay at Home orders

```{r plot-stay-at-home, echo=FALSE, fig.width=8, fig.height = 8}
facetedPlotWithMandates(covidByStateSmoothed, joinMandatesWithCases(mandatesByState, "stayhome", covidByStateSmoothed), "staying at home")
```


## The effect of school closures

```{r plot-school, echo=FALSE, fig.width=8, fig.height = 8}
facetedPlotWithMandates(covidByStateSmoothed, joinMandatesWithCases(mandatesByState, "schools", covidByStateSmoothed), "school closures")
```

```{r plot-nonessential, echo=FALSE, fig.width=8, fig.height = 8}
facetedPlotWithMandates(covidByStateSmoothed, joinMandatesWithCases(mandatesByState, "nonessential", covidByStateSmoothed), "closing non-essential services")
```
